Precise Data Flow Analysis\\in the Presence of Correlated Method Calls\footnote{This research was supported by the Natural Sciences and Engineering Research Council of Canada and the Ontario Ministry of Research and Innovation.}

### 2015-06-19 sas2015_final/content/abstract.tex
creating final version without comments and redundant files

	We show how infeasible paths due to correlated calls can be eliminated for 
	\textit{Interprocedural Finite Distributive Subset} (IFDS) problems, a large class 
	of data-flow analysis problems with broad applications. Our approach is to
	transform an IFDS problem into an \textit{Interprocedural Distributive 
	Environment} (IDE) problem, in which edge functions ``filter out'' data flow 
	along infeasible paths. A solution
	to this IDE problem can be mapped back to the solution space of the original
	IFDS problem.
        We formalize the approach, prove it correct, and report on an implementation
        in the WALA analysis framework.
\end{abstract}

>>>>

	We show how infeasible paths due to correlated calls can be eliminated for 
	\textit{Interprocedural Finite Distributive Subset} (IFDS) problems, a large class 
	of data-flow analysis problems with broad applications. Our approach is to
	transform an IFDS problem into an \textit{Interprocedural Distributive 
	Environment} (IDE) problem, in which edge functions filter out data flow 
	along infeasible paths. A solution
	to this IDE problem can be mapped back to the solution space of the original
	IFDS problem.
        We formalize the approach, prove it correct, and report on an implementation
        in the WALA analysis framework.
\end{abstract}

### 2015-06-19 sas2015_final/content/appendix.tex
creating final version without comments and redundant files

\equalEq*
\begin{proof}
   % OL: Can't we just say the following?
   First, we need to show that for all $r\in R$, if $I(r)=I'(r)$ and $U(r)=U'(r)$, then $\denote{\setpair IU}=\denote{\setpair{I'}{U'}}$. Indeed, we can see that
    \begin{align*}
        \denote{\setpair {I}{U}}
       & = \lambda m. \lambda r. (m(r) \cap I(r))\cup U(r))\\
       & = \lambda m. \lambda r. (m(r) \cap I'(r))\cup U'(r))\\
       & = 
        \denote{\setpair {I'}{U'}}.
    \end{align*}

%OL: Again, we can keep it simpler:
For the other direction:
\begin{align*}
&&
    \denote{\setpair{I}{U}} &= \denote{\setpair{I'}{U'}}\\
&\implies &
    \denote{\setpair{I}{U}}(\lambda r. \emptyset) &= \denote{\setpair{I'}{U'}}(\lambda r. \emptyset)\\
&\implies &
(\lambda m.\lambda r.(m(r)\cap I(r))\cup U(r))(\lambda r.\emptyset) &=
(\lambda m.\lambda r.(m(r)\cap I'(r))\cup U'(r))(\lambda r.\emptyset) \\
&\implies &
\lambda r.(\emptyset \cap I(r))\cup U(r) &=
\lambda r.(\emptyset \cap I'(r))\cup U'(r) \\
&\implies &
\lambda r.U(r) &=
\lambda r.U'(r) \\
&\implies &
U &=
U' \\
\end{align*}
Similarly:
\begin{align*}
&&
    \denote{\setpair{I}{U}} &= \denote{\setpair{I'}{U'}}\\
&\implies &
\denote{\setpair{I}{U}}(I) &= \denote{\setpair{I'}{U'}}(I)\\
&\implies &
(\lambda m.\lambda r.(m(r)\cap I(r))\cup U(r))(I) &=
(\lambda m.\lambda r.(m(r)\cap I'(r))\cup U'(r))(I) \\
&\implies &
\lambda r.(I(r) \cap I(r))\cup U(r) &=
\lambda r.(I(r) \cap I'(r))\cup U'(r) \\
&\implies &
\lambda r.(I(r) \cap I(r)) &=
\lambda r.(I(r) \cap I'(r)) \textrm{ since $\forall r.U(r) \subseteq I(r)$} \\
&\implies \forall r.&
I(r) &= I(r) \cap I'(r)\\
&\implies \forall r.&
I(r) &\subseteq I'(r)\\
\end{align*}
%\mtodo{I agree that this is much better than what I had. I think from $I(r) = I(r) \cap I'(r)$ we can conclude that $I(r)=I'(r)$.}
%\otodo{No. {1,2} = {1,2} \cap {1,2,3}, but {1,2} is not equal to {1,2,3}.}
Symmetrically, we can also establish that $\forall r.I'(r) \subseteq I(r)$ by applying the functions to $I'$ instead of to $I$. Therefore, $I = I'$.
\end{proof}

\equalComp*
%\otodo{The (m)(r) does not seem necessary. Can we replace $r'$ with $r$ and $m'$ with $m$?}
\begin{proof}
\begin{align*}
  &\denote{\setpair IU\circ\setpair{I'}{U'}}\\
  =&\denote{\setpair{\lambda r\,.\,(I(r')\cap I'(r))\cup U(r)}
                    {\lambda r\,.\,(I(r)\cap U'(r))\cup U(r)}}\\
  =&\lambda m\,.\,\lambda r\,.\,(m(r)\cap(I(r)\cap I'(r))\cup U(r))\cup(I(r)\cap U'(r))\cup U(r)\\
  =&\lambda m\,.\,\lambda r\,.\,(m(r)\cap I'(r))\cup I(r)\cup (U'(r)\cap I(r))\cup U(r)\\
  =&\lambda m\,.\,\lambda r\,.\,(((m(r)\cap I'(r))\cup U'(r))\cap I(r))\cup U(r)\\
  =&\lambda m\,.\,\left(\lambda m'\,.\,\lambda r\,.\,(m'(r)\cap I(r))\cup U(r)\right)
    \left(
    \left(\lambda r\,.\,(m(r)\cap I'(r))\cup U'(r)\right)
    \right)\\
  =&\left(\lambda m\,.\,\lambda r\,.\,(m(r)\cap I(r))\cup U(r)\right)\circ
    \left(\lambda m\,.\,\lambda r\,.\,(m(r)\cap I'(r))\cup U'(r)\right)\\
  =&\denote{\setpair IU}\circ\denote{\setpair{I'}{U'}}.\qedhere
\end{align*}
\end{proof}

@@

    Furthermore, $\lcc_\rcc$ is a finite set. 
    Every element of $\lcc_\rcc$ can be applied to $|\rcc|$ receivers, where each receiver is mapped to a set of types. There are $|\rcc|\cdot2^{|T|}$ different possibilities to form those mappings, so
    \[
      |\lcc_\rcc|=|\rcc|\cdot2^{|T|}.
    \]
    Therefore, $\lcc_\rcc\mapsto \lcc_\rcc$ also forms a finite-height lattice.
    \item All operations can be computed in $O(\rcc\times T)$ time. %Note that the $\rcc$ and $T$ sets are an input to the correlated-calls analysis, and the time it takes to compute the meet or composition of micro functions is independent of the representation of the specific operand micro functions.
    \item The space bound is $O(\rcc\times T)$.\qedhere
  \end{enumerate}
\end{proof}

@@

As previously we will denote the top element in the environment lattice as~$\top_\textsf{Env}$.
%\otodo{I've been using $\top_\textsf{Env}$ for clarity.}

@@

As shown in~Sagiv et al.~\cite{sagiv1996precise}, the relationship between environment transformers and edge functions can be described with the following equation. For an edge $(n_1,\,n_2)\in E^*$ an environment $\textsf{env}$ that maps $D$ to $L$, and a fact $d\in D$,
\begin{align}\label{eq:envTransToEdgeFnEdge}
  \menv&((n_1,\,n_2))(\textsf{env})(d)\notag\\
  &=\edgefn((n_1,\,\mathbf0),\,(n_2,\,d))(\top)\sqcap\bigsqcap_{d'\in D}\edgefn((n_1,\,d'),\,(n_2,\,d))(\textsf{env}(d'))\,.
\end{align}  
Then, according to~\eqref{eq:envTransToEdgeFnEdge},
%\mtodo{Is it okay that in~\eqref{eq:envTransToEdgeFnEdge}, we use $\edgefn$, and below we use $\ccedgefn R$?}
%\otodo{Yes, that's fine.}
  \begin{align*}
    &\menv(e_k)\left(\menv(p)(\top_\textsf{Env})\right)(d)(r)\\
      =&\bigg(\ccedgefn R((n_{k-1},\,\mathbf0),\,(n_k,\,d))(\topcc)\sqcap\\
       &\bigsqcap_{d'\in D}\ccedgefn R((n_{k-1},\,d'),\,(n_k,\,d))(\menv(p)(\top_\textsf{Env})(d'))\bigg)(r)\\
      \supseteq&
        \bigsqcap_{d'\in D}\ccedgefn R((n_{k-1},\,d'),\,(n_k,\,d))(\menv(p)(\top_\textsf{Env})(d'))(r)\\
      \supseteq&\,\ccedgefn R((n_{k-1},\,d_{k-1}),\,(n_k,\,d))(\mpdkm)(r)\,.
  \end{align*}
  Therefore, 
  \begin{equation}\label{eq:musubset}
    \efek\subseteq\mppd(r)\,.
  \end{equation}

@@

    The fact that $\mpd(r)\ne\top_T$ means that there exists a realizable path corresponding to the fact $d$ along path $p$, and, by definition, $d$ must be contained in $M_F(p)(\varnothing)$.
    %\mtodo{Is this hand-wavy?}
    %\otodo{No, I think it's fine.}
\end{proof}

@@

\precision*
\begin{proof}
    %\otodo{It's Section 3.2.}
    %\mtodo{Section 3.2 talks about the IFDS algorithm, but it doesn't explicitly say how the result of an IFDS analysis looks like and doesn't introduce the notation $\resultIFDS$.}
  Let $P$ be an IFDS problem. Recall from Section~\ref{sec:ideToIfds} that the result of an IFDS analysis $\resultIFDS(P)$ maps supergraph
  nodes $n\in N^*$ to sets of data-flow facts $\delta\in2^D$. Specifically,
  \begin{align*}
    \resultIFDS(P)&=\lambda n\,.\,\mvp F(n)\\
    &=\bigsqcap_{q\in\textsf{VP}(n)}M_F(q)(\top).
  \end{align*}
  At the same time,
  \begin{align*}
    \backCC\left(\resultIDE(\transCC_R(P))\right)(n)
    &=\left\{d\,|\,\forall r\in R\,.\,\resultIDE(\transCC_R(P))(n)(d)(r)\ne\top_T\right\}\\
    &=\left\{d\,|\,\forall r\in R\,.\,\bigsqcap_{q\in\textsf{VP}(n)}\xi(q,\,d)(r)\ne\top_T\right\}.
  \end{align*}
  This means that for any given $d\in\backCC(\resultIDE(\transCC R(P)))(n)$ and path $q\in\textsf{VP}(n)$, $\xi(q,\,d)\ne\top_T$. Therefore, according to~Lemma~\ref{lem:sound3}, $d\in M_F(q)(\top)$. It follows that $d\in\bigsqcap_{q\in\textsf{VP}(n)}M_F(q)(\top)$, which is equal to $\resultIFDS(P)(n)$. 
  Hence, we have shown that if $d\in\backCC(\resultIDE(\transCC R(P)))(n)$, then $d\in\resultIFDS(P)(n)$.
\end{proof}

@@

%\otodo{restate the lemma before the proof}
\onlycorrrec*
\begin{proof}[\textbf{Proof of Lemma~\ref{lem:onlycorrrec}}]
  By definition of $\backCC$,
  \begin{align*}
    \backCC(\resultIDE\left(\transCC_R(P)\right))
    &=\{d\,|\,\resultIDE(\transCC_R(P))(n)(d)=\ell,\,\forall r\,.\,\ell(r)\ne\top_T\}\\
    &=\{d\,|\,\forall r\in R,\,\resultIDE(\transCC_R(P))(n)(d)(r)\ne\top_T\}.
  \end{align*}

  \commentout{
  \begin{align*}
    \backCC(\resultIDE\left(\transCC_R(P)\right))
    &=\left\{(n,\,D_n^\Subset(\resultIDE(\transCC_R(P))))\ |\ n\in N^*\right\}.
  \end{align*}
  According to~\eqref{eq:dnq} and~Lemma~\ref{lem:recindep}, for a given $n\in N^*$,
  \begin{align*}
    D_n^\Subset&(\resultIDE(\transCC_R(P))))\\
    &=\left\{d\ |\ d\in\mvp F(n)\,\wedge\,\forall r\in R:\ \left\{(r,\,\resultIDE(\transCC_{\{r\}}(P))(n,\,d)(r))\ |\ r\in R\right\}(r)\ne\top_T\right\}\\
    &=\left\{d\ |\ d\in\mvp F(n)\,\wedge\,\forall\bm{r\in R}:\ \resultIDE(\transCC_{\{r\}}(P))(n,\,d)(r)\ne\top_T\right\}.
  \end{align*}
  }
Since, according to Lemma~\ref{lem:ccrectop}, $\resultIDE(\transCC_{\{r\}}(P))(n)(d)(r)$ can only be equal to $\top_T$ when $r\in\rcc$, we can conclude that
  \begin{align*}
    \backCC(\resultIDE\left(\transCC_R(P)\right))
    &=\{d\,|\,\bm{\forall r\in\rcc},\,\resultIDE(\transCC_{\bm\rcc}(P))(n)(d)(r)\ne\top_T\}\\
    &=\backCC(\resultIDE\left(\transCC_\rcc(P)\right)).
    \qedhere
  \end{align*}
\end{proof}

>>>>

\equalEq*
\begin{proof}
   First, we need to show that for all $r\in R$, if $I(r)=I'(r)$ and $U(r)=U'(r)$, then $\denote{\setpair IU}=\denote{\setpair{I'}{U'}}$. Indeed, we can see that
    \begin{align*}
        \denote{\setpair {I}{U}}
       & = \lambda m. \lambda r. (m(r) \cap I(r))\cup U(r))\\
       & = \lambda m. \lambda r. (m(r) \cap I'(r))\cup U'(r))\\
       & = 
        \denote{\setpair {I'}{U'}}.
    \end{align*}

For the other direction:
\begin{align*}
&&
    \denote{\setpair{I}{U}} &= \denote{\setpair{I'}{U'}}\\
&\implies &
    \denote{\setpair{I}{U}}(\lambda r. \emptyset) &= \denote{\setpair{I'}{U'}}(\lambda r. \emptyset)\\
&\implies &
(\lambda m.\lambda r.(m(r)\cap I(r))\cup U(r))(\lambda r.\emptyset) &=
(\lambda m.\lambda r.(m(r)\cap I'(r))\cup U'(r))(\lambda r.\emptyset) \\
&\implies &
\lambda r.(\emptyset \cap I(r))\cup U(r) &=
\lambda r.(\emptyset \cap I'(r))\cup U'(r) \\
&\implies &
\lambda r.U(r) &=
\lambda r.U'(r) \\
&\implies &
U &=
U' \\
\end{align*}
Similarly:
\begin{align*}
&&
    \denote{\setpair{I}{U}} &= \denote{\setpair{I'}{U'}}\\
&\implies &
\denote{\setpair{I}{U}}(I) &= \denote{\setpair{I'}{U'}}(I)\\
&\implies &
(\lambda m.\lambda r.(m(r)\cap I(r))\cup U(r))(I) &=
(\lambda m.\lambda r.(m(r)\cap I'(r))\cup U'(r))(I) \\
&\implies &
\lambda r.(I(r) \cap I(r))\cup U(r) &=
\lambda r.(I(r) \cap I'(r))\cup U'(r) \\
&\implies &
\lambda r.(I(r) \cap I(r)) &=
\lambda r.(I(r) \cap I'(r)) \textrm{ since $\forall r.U(r) \subseteq I(r)$} \\
&\implies \forall r.&
I(r) &= I(r) \cap I'(r)\\
&\implies \forall r.&
I(r) &\subseteq I'(r)\\
\end{align*}
Symmetrically, we can also establish that $\forall r.I'(r) \subseteq I(r)$ by applying the functions to $I'$ instead of to $I$. Therefore, $I = I'$.
\end{proof}

\equalComp*
\begin{proof}
\begin{align*}
  &\denote{\setpair IU\circ\setpair{I'}{U'}}\\
  =&\denote{\setpair{\lambda r\,.\,(I(r')\cap I'(r))\cup U(r)}
                    {\lambda r\,.\,(I(r)\cap U'(r))\cup U(r)}}\\
  =&\lambda m\,.\,\lambda r\,.\,(m(r)\cap(I(r)\cap I'(r))\cup U(r))\cup(I(r)\cap U'(r))\cup U(r)\\
  =&\lambda m\,.\,\lambda r\,.\,(m(r)\cap I'(r))\cup I(r)\cup (U'(r)\cap I(r))\cup U(r)\\
  =&\lambda m\,.\,\lambda r\,.\,(((m(r)\cap I'(r))\cup U'(r))\cap I(r))\cup U(r)\\
  =&\lambda m\,.\,\left(\lambda m'\,.\,\lambda r\,.\,(m'(r)\cap I(r))\cup U(r)\right)
    \left(
    \left(\lambda r\,.\,(m(r)\cap I'(r))\cup U'(r)\right)
    \right)\\
  =&\left(\lambda m\,.\,\lambda r\,.\,(m(r)\cap I(r))\cup U(r)\right)\circ
    \left(\lambda m\,.\,\lambda r\,.\,(m(r)\cap I'(r))\cup U'(r)\right)\\
  =&\denote{\setpair IU}\circ\denote{\setpair{I'}{U'}}.\qedhere
\end{align*}
\end{proof}

@@

    Furthermore, $\lcc_\rcc$ is a finite set. 
    Every element of $\lcc_\rcc$ can be applied to $|\rcc|$ receivers, where each receiver is mapped to a set of types. There are $|\rcc|\cdot2^{|T|}$ different possibilities to form those mappings, so
    \[
      |\lcc_\rcc|=|\rcc|\cdot2^{|T|}.
    \]
    Therefore, $\lcc_\rcc\mapsto \lcc_\rcc$ also forms a finite-height lattice.
    \item All operations can be computed in $O(\rcc\times T)$ time.
    \item The space bound is $O(\rcc\times T)$.\qedhere
  \end{enumerate}
\end{proof}

@@

As previously we will denote the top element in the environment lattice as~$\top_\textsf{Env}$.

@@

As shown in~Sagiv et al.~\cite{sagiv1996precise}, the relationship between environment transformers and edge functions can be described with the following equation. For an edge $(n_1,\,n_2)\in E^*$ an environment $\textsf{env}$ that maps $D$ to $L$, and a fact $d\in D$,
\begin{align}\label{eq:envTransToEdgeFnEdge}
  \menv&((n_1,\,n_2))(\textsf{env})(d)\notag\\
  &=\edgefn((n_1,\,\mathbf0),\,(n_2,\,d))(\top)\sqcap\bigsqcap_{d'\in D}\edgefn((n_1,\,d'),\,(n_2,\,d))(\textsf{env}(d'))\,.
\end{align}  
Then, according to~\eqref{eq:envTransToEdgeFnEdge},
  \begin{align*}
    &\menv(e_k)\left(\menv(p)(\top_\textsf{Env})\right)(d)(r)\\
      =&\bigg(\ccedgefn R((n_{k-1},\,\mathbf0),\,(n_k,\,d))(\topcc)\sqcap\\
       &\bigsqcap_{d'\in D}\ccedgefn R((n_{k-1},\,d'),\,(n_k,\,d))(\menv(p)(\top_\textsf{Env})(d'))\bigg)(r)\\
      \supseteq&
        \bigsqcap_{d'\in D}\ccedgefn R((n_{k-1},\,d'),\,(n_k,\,d))(\menv(p)(\top_\textsf{Env})(d'))(r)\\
      \supseteq&\,\ccedgefn R((n_{k-1},\,d_{k-1}),\,(n_k,\,d))(\mpdkm)(r)\,.
  \end{align*}
  Therefore, 
  \begin{equation}\label{eq:musubset}
    \efek\subseteq\mppd(r)\,.
  \end{equation}

@@

    The fact that $\mpd(r)\ne\top_T$ means that there exists a realizable path corresponding to the fact $d$ along path $p$, and, by definition, $d$ must be contained in $M_F(p)(\varnothing)$.
\end{proof}

@@

\precision*
\begin{proof}
  Let $P$ be an IFDS problem. Recall from Section~\ref{sec:ideToIfds} that the result of an IFDS analysis $\resultIFDS(P)$ maps supergraph
  nodes $n\in N^*$ to sets of data-flow facts $\delta\in2^D$. Specifically,
  \begin{align*}
    \resultIFDS(P)&=\lambda n\,.\,\mvp F(n)\\
    &=\bigsqcap_{q\in\textsf{VP}(n)}M_F(q)(\top).
  \end{align*}
  At the same time,
  \begin{align*}
    \backCC\left(\resultIDE(\transCC_R(P))\right)(n)
    &=\left\{d\,|\,\forall r\in R\,.\,\resultIDE(\transCC_R(P))(n)(d)(r)\ne\top_T\right\}\\
    &=\left\{d\,|\,\forall r\in R\,.\,\bigsqcap_{q\in\textsf{VP}(n)}\xi(q,\,d)(r)\ne\top_T\right\}.
  \end{align*}
  This means that for any given $d\in\backCC(\resultIDE(\transCC R(P)))(n)$ and path $q\in\textsf{VP}(n)$, $\xi(q,\,d)\ne\top_T$. Therefore, according to~Lemma~\ref{lem:sound3}, $d\in M_F(q)(\top)$. It follows that $d\in\bigsqcap_{q\in\textsf{VP}(n)}M_F(q)(\top)$, which is equal to $\resultIFDS(P)(n)$. 
  Hence, we have shown that if $d\in\backCC(\resultIDE(\transCC R(P)))(n)$, then $d\in\resultIFDS(P)(n)$.
\end{proof}

@@

\onlycorrrec*
\begin{proof}[\textbf{Proof of Lemma~\ref{lem:onlycorrrec}}]
  By definition of $\backCC$,
  \begin{align*}
    \backCC(\resultIDE\left(\transCC_R(P)\right))
    &=\{d\,|\,\resultIDE(\transCC_R(P))(n)(d)=\ell,\,\forall r\,.\,\ell(r)\ne\top_T\}\\
    &=\{d\,|\,\forall r\in R,\,\resultIDE(\transCC_R(P))(n)(d)(r)\ne\top_T\}.
  \end{align*}

Since, according to Lemma~\ref{lem:ccrectop}, $\resultIDE(\transCC_{\{r\}}(P))(n)(d)(r)$ can only be equal to $\top_T$ when $r\in\rcc$, we can conclude that
  \begin{align*}
    \backCC(\resultIDE\left(\transCC_R(P)\right))
    &=\{d\,|\,\bm{\forall r\in\rcc},\,\resultIDE(\transCC_{\bm\rcc}(P))(n)(d)(r)\ne\top_T\}\\
    &=\backCC(\resultIDE\left(\transCC_\rcc(P)\right)).
    \qedhere
  \end{align*}
\end{proof}

### 2015-06-19 sas2015_final/content/background.tex
creating final version without comments and redundant files

\section{Background}\label{sec:bg}
%The purpose of the correlated-calls analysis is to solve IFDS problems more precisely than using the standard IFDS algorithm by ruling out some infeasible paths. The correlated-calls analysis works by transforming an IFDS problem to an IDE problem, solving the IDE problem, and transforming the IDE result to a solution to the original IFDS problem. 
This section defines terminology and presents the IFDS and IDE algorithms.
%\ftodo{
  %Shall we add a sentence saying that readers familiar with these algorithms may skip this section?
%}

\subsection{Terminology and Notation}
%We will start by introducing several concepts used by the IFDS and IDE analyses.
%\begin{mdelete}\mtodo{We already say (but less precisely) what a CFG is in the introduction.}
The \textit{control-flow graph} of a procedure is a directed graph whose nodes are instructions, which contains an edge from $n_1$ to $n_2$ whenever
$n_2$ may execute immediately after $n_1$. A CFG has a distinguished \textit{start node} \textsf{start}$_p$ and \textit{end node} \textsf{end}$_p$.
Following the presentation of Reps~et~al.~\cite{reps1995precise,sagiv1996precise}, we follow every
call instruction with a no-op instruction, so that every \textit{call node} is immediately
followed by a \textit{return node} in the CFG.
The \textit{control-flow supergraph} of a program contains the CFGs of all of the procedures as
subgraphs. In addition, for each call instruction $c$, the supergraph contains a \textit{call-to-start} edge to the start node of every procedure that
may be called from $c$, and an \textit{end-to-return} edge from the end node of the procedure back to the call instruction.

@@

A path in the control-flow supergraph is \textit{valid} if it follows the usual stack-based calling 
discipline: every end-to-return edge on the path returns
to the site of the most recent call that has not yet been matched by a return. 
%\ftodo{Can we explain/define the notion of the ``most recent unmatched call'' more precisely?}
The set of all valid paths from the program entry point to
a node $n$ is denoted $\ivp(n)$.

@@

\subsection{IFDS}\label{sec:bgifds}
The IFDS framework~\cite{reps1995precise} is a precise and efficient algorithm for data-flow analysis
%. IFDS was developed in 1995 by T.\,Reps, S.\,Horwitz, and M.\,Sagiv at the University of Wisconsin and
that has been used to solve a variety of data-flow analysis problems~\cite{bodden2013spl,naeem2008typestate,DBLP:conf/birthday/KreikerRRSWY13,tripp2009taj}.
The IFDS framework is an instance of the \textit{functional approach} to data-flow analysis~\cite{pnueli1981two}
because it constructs summaries of the effects of called procedures.
The IFDS framework is applicable to \textit{interprocedural} data-flow problems whose domain consists of \textit{subsets} of a \textit{finite} set $D$,
and whose data-flow functions are \textit{distributive}.
A function $f$ is distributive if 
$f(x_1\sqcap x_2)=f(x_1)\sqcap f(x_2)$.

@@

The input to the IFDS algorithm is specified as
%Formally, a data-flow analysis problem is suitable for an IFDS analysis if it can be encoded as an IFDS problem
  $(G^*,\,D,\,F,\,M_F,\,\sqcap)$,
where $G^*=(N^*,\,E^*)$ is the supergraph of the input program with nodes $N^*$ and edges $E^*$,
$D$ is a finite set of \textit{data-flow facts},
$F$ is a set of distributive data-flow functions of type $2^D\to2^D$,
$M_F:\,E^*\to F$ assigns a data-flow function to each supergraph edge,
and $\sqcap$ is the \textit{meet operator} on the powerset $2^D$, either union or intersection.
In our presentation, the meet operator will always be union, but all of the results apply dually when the
meet is intersection.

@@

The output of the IDE algorithm is, for each node $n$ in the supergraph,
the \textit{meet-over-all-valid-paths} solution
$\mvp{\textsf{Env}}(n)=\bigsqcap_{q\in\ivp(n)}M_\textsf{Env}(q)(\top_\textsf{Env})$,
%\mtodo{Should it be $\Omega$ instead of $\top$?}
where $\top_\textsf{Env} = \lambda d.\top$ is the top element of the lattice of environments,
and $M_\textsf{Env}$ is extended from edges to paths by composition.

@@

A practical implementation of the IDE algorithm requires
the input data-flow functions to be provided in their pointwise representation
as exploded supergraph edges labelled with micro-functions.
%\mtodo{Should we cite the ``Practical extensions'' paper here that talks about the on-demand IFDS algorithm?}
Specifically,
the input is generally provided as a function $\textsf{EdgeFn}: (N^* \times D) \times (N^* \times D) \to F$,
where $F$ is the set of representations of micro-functions from $L$ to $L$.
Given an exploded supergraph edge $e=(n,\,d) \to (n',\,d')$,
$\textsf{EdgeFn}((n,\,d),\,(n',\,d'))$ returns the micro-function that appears on the exploded supergraph edge $e$.
%Like in the IFDS algorithm,
In an implementation, it can be convenient to split the function $\textsf{EdgeFn}$ into separate functions that handle the cases when $n\to n'$ is an intraprocedural edge, a call-to-return edge, a call-to-start edge, or an end-to-return edge.

>>>>

\section{Background}\label{sec:bg}
This section defines terminology and presents the IFDS and IDE algorithms.

\subsection{Terminology and Notation}
The \textit{control-flow graph} of a procedure is a directed graph whose nodes are instructions, which contains an edge from $n_1$ to $n_2$ whenever
$n_2$ may execute immediately after $n_1$. A CFG has a distinguished \textit{start node} \textsf{start}$_p$ and \textit{end node} \textsf{end}$_p$.
Following the presentation of Reps~et~al.~\cite{reps1995precise,sagiv1996precise}, we follow every
call instruction with a no-op instruction, so that every \textit{call node} is immediately
followed by a \textit{return node} in the CFG.
The \textit{control-flow supergraph} of a program contains the CFGs of all of the procedures as
subgraphs. In addition, for each call instruction $c$, the supergraph contains a \textit{call-to-start} edge to the start node of every procedure that
may be called from $c$, and an \textit{end-to-return} edge from the end node of the procedure back to the call instruction.

@@

A path in the control-flow supergraph is \textit{valid} if it follows the usual stack-based calling 
discipline: every end-to-return edge on the path returns
to the site of the most recent call that has not yet been matched by a return. 
The set of all valid paths from the program entry point to
a node $n$ is denoted $\ivp(n)$.

@@

\subsection{IFDS}\label{sec:bgifds}
The IFDS framework~\cite{reps1995precise} is a precise and efficient algorithm for data-flow analysis
that has been used to solve a variety of data-flow analysis problems~\cite{bodden2013spl,naeem2008typestate,DBLP:conf/birthday/KreikerRRSWY13,tripp2009taj}.
The IFDS framework is an instance of the \textit{functional approach} to data-flow analysis~\cite{pnueli1981two}
because it constructs summaries of the effects of called procedures.
The IFDS framework is applicable to \textit{interprocedural} data-flow problems whose domain consists of \textit{subsets} of a \textit{finite} set $D$,
and whose data-flow functions are \textit{distributive}.
A function $f$ is distributive if 
$f(x_1\sqcap x_2)=f(x_1)\sqcap f(x_2)$.

@@

The input to the IFDS algorithm is specified as
  $(G^*,\,D,\,F,\,M_F,\,\sqcap)$,
where $G^*=(N^*,\,E^*)$ is the supergraph of the input program with nodes $N^*$ and edges $E^*$,
$D$ is a finite set of \textit{data-flow facts},
$F$ is a set of distributive data-flow functions of type $2^D\to2^D$,
$M_F:\,E^*\to F$ assigns a data-flow function to each supergraph edge,
and $\sqcap$ is the \textit{meet operator} on the powerset $2^D$, either union or intersection.
In our presentation, the meet operator will always be union, but all of the results apply dually when the
meet is intersection.

@@

The output of the IDE algorithm is, for each node $n$ in the supergraph,
the \textit{meet-over-all-valid-paths} solution
$\mvp{\textsf{Env}}(n)=\bigsqcap_{q\in\ivp(n)}M_\textsf{Env}(q)(\top_\textsf{Env})$,
where $\top_\textsf{Env} = \lambda d.\top$ is the top element of the lattice of environments,
and $M_\textsf{Env}$ is extended from edges to paths by composition.

@@

A practical implementation of the IDE algorithm requires
the input data-flow functions to be provided in their pointwise representation
as exploded supergraph edges labelled with micro-functions.
Specifically,
the input is generally provided as a function $\textsf{EdgeFn}: (N^* \times D) \times (N^* \times D) \to F$,
where $F$ is the set of representations of micro-functions from $L$ to $L$.
Given an exploded supergraph edge $e=(n,\,d) \to (n',\,d')$,
$\textsf{EdgeFn}((n,\,d),\,(n',\,d'))$ returns the micro-function that appears on the exploded supergraph edge $e$.
In an implementation, it can be convenient to split the function $\textsf{EdgeFn}$ into separate functions that handle the cases when $n\to n'$ is an intraprocedural edge, a call-to-return edge, a call-to-start edge, or an end-to-return edge.
